---
title: "NAITRE Trial CONSORT Participant Flow"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: 
- Fanice Nyatigo fanice.nyatigo@ucsf.edu
- Ben Arnold ben.arnold@ucsf.edu 
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---

# Summary

Track participant flow in the NAITRE trial.

Summarize patient numbers by group to include in a CONSORT flow diagram.

# Configuration

```{r preamble, message = FALSE}
#---------------------------------
# source the project's configuration
# file
#---------------------------------
library(here)
source(here("NAITRE-primary/R","NAITRE-primary-Config.R"))
```


# Load the data

Load the final population tracking dataset. This file was created by `0-NAITRE-primary-data-processing.R`.

```{r load the data}
#---------------------------------
# load the preprocessed data
#---------------------------------

# consort data
df_consort <- read_rds(here("NAITRE-primary/data/final","NAITRE_consort_public.rds"))

# primary analysis file
df_analysis <- read_rds(here("NAITRE-primary/data/final","NAITRE_primary_analysis_public.rds"))

```

# CONSORT flow

Participant flow by study visit

## Screening and Enrollment
```{r consort flow screening}
#---------------------------------
# Screening and Enrollment
#---------------------------------
dc1 <- df_consort %>% summarise(
  `Assessed for eligibility`                       = sum(screened == 1),
  `Excluded (Overall) *`                           = sum(enrolled != 1 | is.na(enrolled)),
  `Under 8 days old`                               = sum(too_young == 1, na.rm = T),
  `Over 27 days old`                               = sum(too_old == 1, na.rm = T),
  `Less than 2500g`                                = sum(ex_weight == 1, na.rm = T),
  `Cannot orally feed`                             = sum(ex_oral == 1, na.rm = T),
  `Will not be in the study area for all 6 months` = sum(ex_stay6months == 1, na.rm = T),
  `Allergy`                                        = sum(ex_allergy == 1, na.rm = T),
  `Participant in another study`                   = sum(ex_long == 1, na.rm = T),
  `Lives outside study areas`                      = sum(ex_village == 1, na.rm = T),
  `Ineligibility reason unknown`                   = sum(unknown == 1, na.rm = T),
  `Refused`                                        = sum(refused == 1, na.rm = T), 
  `Enrolled`                                       = sum(enrolled == 1, na.rm = T)
) %>% 
  # transpose
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "Variable") %>% 
  rename(Number = V1) 

kable(dc1) %>% 
  row_spec(c(1, 2, 13), bold = T, italic = T)
```
*Note - the reasons for ineligibility are not mutually exclusive; a child may have more than one reason for exclusion.

## Allocation and Follow-up
```{r consort flow followup}
#---------------------------------
# among participants enrolled
# identify treatment and 
# participants lost to follow-up 
# at various points
#---------------------------------
dc2 <- df_consort %>%
  # restrict to participants enrolled
  filter(enrolled == 1) %>%
         # treatment administration
  mutate(treated = case_when(isAdministered == 1 ~ "Yes",
                             isAdministered == 2 ~ "No",
                             is.na(isAdministered) ~ NA_character_),
         # analyzed for AE?
         an_ae = factor(an_ae, levels=c(1,0), labels = c("Yes","No")),)

#---------------------------------
# label the variables for table
# printing
#---------------------------------

label(dc2$treated) <- "Treatment Administered"
label(dc2$an_ae) <- "Assessed for Adverse Events"
label(dc2$d_21_status) <- "Status, within first 21 days of enrollment"
label(dc2$d_90_status) <- "Status, 3 months since birth"
label(dc2$d_180_status) <- "Status, 6 months since birth"

table1(~ treated + an_ae + d_21_status + d_90_status + d_180_status | tx, data = dc2)

```

# Age distribution at 6-month followup

```{r}
ggplot(
  df_analysis %>% select(age_days_180), aes(x = age_days_180)) + 
  geom_histogram(alpha = 1, binwidth = 1) +
  geom_vline(xintercept = c(141, 225), color="red", linetype = "dashed") +
  geom_vline(xintercept = 180, color="black", linetype = "dashed") +
  labs(x = "Age in days", y= "Count", 
       caption = "*Red lines represent age limit boundaries") +
  geom_text(aes(x = 180, y = 1000, label = "180 days"))+
       theme_bw()
```

# Session Info
```{r session info}
sessionInfo()
```
