---
title: "NAITRE Trial Baseline Balance"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Fanice Nyatigo, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---



# Summary

This report includes unmasked results from the NAITRE trial ([NCT03682653](https://clinicaltrials.gov/ct2/show/NCT03682653)). 

The report summarizes baseline characteristics by group

Should you have any questions regarding this report, please write Dr. Benjamin Arnold (ben.arnold@ucsf.edu).

# Preamble
```{r preamble, message = FALSE}
#------------------------------
# load configuration file
# includes packages and masked
# assignments
#------------------------------
library(here)
source(here("NAITRE-primary/R","NAITRE-primary-Config.R"))

```


# Load the data

```{r read in clean data} 
#---------------------------------
# load the preprocessed data
#---------------------------------

# primary analysis file
df_analysis <- read_rds(here("NAITRE-primary/data/final","NAITRE_primary_analysis_public.rds"))

```


# Baseline characteristics by group

```{r baseline indicators}
# custom render function for continuous variables - just show MEAN and SD
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=1), c("",
        "Mean (SD)"=sprintf("%s (%s)", MEAN, SD)))
}


# prepare variable for rendering table
df_analysis %>%
  mutate(Sex = childSex,
         Region = factor(region_name, levels = c("Centre", "Boucle du Mouhoun", "Cascade", 
                                                   "Centre Ouest", "Hauts-Bassins")),
         `Weight at enrollment (g)` = weight_0*1000,
         `Season of enrollment` = enroll_season,
         `Pregnancy type` = case_when(preg_type == 1 ~ "Singleton",
                                      preg_type == 2 ~ "Multiple",
                                      is.na(preg_type) ~ NA_character_),
         `Pregnancy type` = factor(`Pregnancy type`, levels = c("Singleton", "Multiple")),
         Breastfeeding = breastfeeding,
         `Inititation of breastfeeding` = case_when(breastfeedInit2 == 1 ~ "Immediate",
                                                    breastfeedInit2 == 2 ~ "Delayed",
                                                    childFeeding == 2 ~ "Does not breastfeed",
                                                    is.na(breastfeedInit2) ~ NA_character_),
         `Inititation of breastfeeding` = factor(`Inititation of breastfeeding`, levels = c("Immediate", "Delayed", "Does not breastfeed")),
         `Mother's age` = mum_age,
         `Mother's education` = case_when(mum_edu == 1 ~ "None",
                                          mum_edu == 2 ~ "Primary",
                                          mum_edu %in% c(3,4) ~ "Secondary or above",
                                          is.na(mum_edu) ~ NA_character_),
         `No. of antenatal visits` = prenatalVisits,
         tx_table1 = factor(tx, levels = c("Azithromycin","Placebo"))
         ) %>%
  set_variable_labels(age_days_0 = "Age (days)", 
                      birthweight = "Birthweight (g)", 
                      length_0 = "Length at enrollment (cm)",
                      ) %>%
  table1(~ age_days_0 + Sex + Region + birthweight + `Weight at enrollment (g)` + `Season of enrollment` + 
           length_0 + `Pregnancy type` + Breastfeeding + `Inititation of breastfeeding` + 
           `Mother's age` + `Mother's education` + `No. of antenatal visits` | tx_table1, data = .,
         render.continuous= c(.="Median (Q1 to Q3)")
         )

```
# Baseline characteristics - included in primary analysis

```{r}
# filter to those who were not ltfu
df_analysis %>% 
  filter(death_status_180 %in% c("Measured, in-window", "Event prior to window")) %>% 
  mutate(Sex = childSex,
          Region = factor(region_name, levels = c("Centre", "Boucle du Mouhoun", "Cascade", 
                                                   "Centre Ouest", "Hauts-Bassins")),
         `Weight at enrollment (g)` = weight_0*1000,
         `Season of enrollment` = enroll_season,
         `Pregnancy type` = case_when(preg_type == 1 ~ "Singleton",
                                      preg_type == 2 ~ "Multiple",
                                      is.na(preg_type) ~ NA_character_),
         `Pregnancy type` = factor(`Pregnancy type`, levels = c("Singleton", "Multiple")),
         Breastfeeding = breastfeeding,
         `Inititation of breastfeeding` = case_when(breastfeedInit2 == 1 ~ "Immediate",
                                                    breastfeedInit2 == 2 ~ "Delayed",
                                                    childFeeding == 2 ~ "Does not breastfeed",
                                                    is.na(breastfeedInit2) ~ NA_character_),
         `Inititation of breastfeeding` = factor(`Inititation of breastfeeding`, levels = c("Immediate", "Delayed", "Does not breastfeed")),
         `Mother's age` = mum_age,
         `Mother's education` = case_when(mum_edu == 1 ~ "None",
                                          mum_edu == 2 ~ "Primary",
                                          mum_edu %in% c(3,4) ~ "Secondary or above",
                                          is.na(mum_edu) ~ NA_character_),
         `No. of antenatal visits` = prenatalVisits,
         tx_table1 = factor(tx, levels = c("Azithromycin","Placebo"))
         ) %>%
  set_variable_labels(age_days_0 = "Age (days)", 
                      birthweight = "Birthweight (g)", 
                      length_0 = "Length at enrollment (cm)",
                      ) %>%
  table1(~ age_days_0 + Sex + Region + birthweight + `Weight at enrollment (g)` + `Season of enrollment` + 
           length_0 + `Pregnancy type` + Breastfeeding + `Inititation of breastfeeding` + 
           `Mother's age` + `Mother's education` + `No. of antenatal visits` | tx_table1, data = .,
         render.continuous= c(.="Median (Q1 to Q3)")
         )
```

# Baseline characteristics - children excluded from primary analysis (measured out of window or lost to followup)
```{r}
# filter to those who were ltfu
df_analysis %>% 
  filter(death_status_180 %in% c("Measured, out-of-window", "Lost to follow-up")) %>% 
  mutate(Sex = childSex,
          Region = factor(region_name, levels = c("Centre", "Boucle du Mouhoun", "Cascade", 
                                                   "Centre Ouest", "Hauts-Bassins")),
         `Weight at enrollment (g)` = weight_0*1000,
         `Season of enrollment` = enroll_season,
         `Pregnancy type` = case_when(preg_type == 1 ~ "Singleton",
                                      preg_type == 2 ~ "Multiple",
                                      is.na(preg_type) ~ NA_character_),
         `Pregnancy type` = factor(`Pregnancy type`, levels = c("Singleton", "Multiple")),
         Breastfeeding = breastfeeding,
         `Inititation of breastfeeding` = case_when(breastfeedInit2 == 1 ~ "Immediate",
                                                    breastfeedInit2 == 2 ~ "Delayed",
                                                    childFeeding == 2 ~ "Does not breastfeed",
                                                    is.na(breastfeedInit2) ~ NA_character_),
         `Inititation of breastfeeding` = factor(`Inititation of breastfeeding`, levels = c("Immediate", "Delayed", "Does not breastfeed")),
         `Mother's age` = mum_age,
         `Mother's education` = case_when(mum_edu == 1 ~ "None",
                                          mum_edu == 2 ~ "Primary",
                                          mum_edu %in% c(3,4) ~ "Secondary or above",
                                          is.na(mum_edu) ~ NA_character_),
         `No. of antenatal visits` = prenatalVisits,
         tx_table1 = factor(tx, levels = c("Azithromycin","Placebo"))
         ) %>%
  set_variable_labels(age_days_0 = "Age (days)", 
                      birthweight = "Birthweight (g)", 
                      length_0 = "Length at enrollment (cm)",
                      ) %>%
  table1(~ age_days_0 + Sex + Region + birthweight + `Weight at enrollment (g)` + `Season of enrollment` +
           length_0 + `Pregnancy type` + Breastfeeding + `Inititation of breastfeeding` + 
           `Mother's age` + `Mother's education` + `No. of antenatal visits` | tx_table1, data = .,
         render.continuous= c(.="Median (Q1 to Q3)")
         )
```

# Session Info
```{r session info}
sessionInfo()
```

