---
title: "NAITRE Trial Adverse Events Analysis"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Fanice Nyatigo, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---

# Summary

This report includes unmasked results from the NAITRE trial ([NCT03682653](https://clinicaltrials.gov/ct2/show/NCT03682653)). 

The report summarizes adverse events by day 21 in the trial.

Should you have any questions regarding this report, please write Dr. Benjamin Arnold (ben.arnold@ucsf.edu).

# Preamble
```{r preamble, message=FALSE}
#------------------------------
# load configuration file
# includes packages and masked
# assignments
#------------------------------
library(here)
source(here("NAITRE-primary/R","NAITRE-primary-Config.R"))

```

# Load the analysis data

```{r read in clean data} 
#------------------------------
# read in the clean data 
# rename public ID to childID
#------------------------------

#  primary analysis file
df_analysis <- read_rds(here("NAITRE-primary/data/final","NAITRE_primary_analysis_public.rds")) %>% 
  rename(childID = id_public)

```


# Adverse Events by 21 days

Tabulation of any ocurrence of each symptom by 21 days among study participants.

```{r SAEs at 21 days}
#-------------------------------
# make the SAE variables factor
# variables for a table
#-------------------------------
d_ae_table <- df_analysis %>%
  mutate(
    ae_fever_21 = factor(ae_fever_21, levels=c(0,1), labels = c("No","Yes")),
    ae_abpain_21 = factor(ae_abpain_21, levels=c(0,1), labels = c("No","Yes")),
    ae_constip_21 = factor(ae_constip_21, levels=c(0,1), labels = c("No","Yes")),
    ae_rash_21 = factor(ae_rash_21, levels=c(0,1), labels = c("No","Yes")),
    ae_diarrhea_21 = factor(ae_diarrhea_21, levels=c(0,1), labels = c("No","Yes")),
    ae_vomiting_21 = factor(ae_vomiting_21, levels=c(0,1), labels = c("No","Yes")),
    ae_vomitAfterfeeding_21 = factor(ae_vomitAfterfeeding_21, levels=c(0,1), labels = c("No","Yes")),
    ae_projVomit_21 = factor(ae_projVomit_21, levels=c(0,1), labels = c("No","Yes")),
    ae_non_serious_21 = factor(ae_non_serious_21, levels=c(0,1), labels = c("No","Yes")),
    ae_ihps_28 = factor(ae_ihps_28, levels=c(0,1), labels = c("No","Yes")),
    ae_death_28 = factor(ae_death_28, levels=c(0,1), labels = c("No","Yes")),
    ae_hosp_28 = factor(ae_hosp_28, levels=c(0,1), labels = c("No","Yes")),
    ae_serious_28 = factor(ae_serious_28, levels=c(0,1), labels = c("No","Yes")),
    tx = factor(tx, levels = c("Azithromycin", "Placebo"))
         )
  
d_ae_table %<>%
  set_variable_labels(ae_fever_21 = "Fever",
                      ae_abpain_21 = "Abdominal pain",
                      ae_constip_21 = "Constipation",
                      ae_rash_21 = "Rash",
                      ae_diarrhea_21 = "Diarrhea",
                      ae_vomiting_21 = "Vomiting (Any)",
                      ae_vomitAfterfeeding_21 = "Vomiting after every feed",
                      ae_projVomit_21 = "Projectile vomiting",
                      ae_non_serious_21 = "Any non-serious adverse event",
                      ae_ihps_28 = "Infantile hypertrophic pyloric stenosis",
                      ae_death_28 = "Mortality within 28 days of treatment",
                      ae_hosp_28 = "Hospitalization within 28 days of treatment",
                      ae_serious_28 = "Any serious adverse event"
                      ) 
```

## Tabulation of SAEs

```{r}
d_ae_table %>% 
  table1(~ ae_serious_28 + ae_ihps_28 + ae_death_28 + ae_hosp_28 | tx, data = .)
```

\vspace{5em}

## Tabulation of non-serious AEs

```{r AEs at 21 days}
d_ae_table %>% 
  table1(~ ae_non_serious_21 + ae_vomiting_21 + ae_vomitAfterfeeding_21 + ae_projVomit_21 + ae_diarrhea_21 + ae_fever_21 + ae_abpain_21 +  ae_rash_21 + ae_constip_21 | tx, data = .)
```

## Differences between groups in SAEs

```{r SAE differences between arms}
#--------------------------
# pivot the AE symptoms to
# long format
#--------------------------
d_ael <- df_analysis %>% 
  select(-c(ae_non_serious_21, ae_vomiting_21, ae_vomitAfterfeeding_21, 
            ae_projVomit_21, ae_diarrhea_21, ae_fever_21, ae_abpain_21, 
            ae_rash_21, ae_constip_21)) %>% 
  group_by(childID,tx) %>%
  select(childID,tx,starts_with("ae")) %>%
  # reshape longer to make it easier to iterate the analysis
  pivot_longer(cols=starts_with("ae_"),names_to = "ae_name",values_to = "ae") %>%
  # ensure treatment assignment is a factor
  # order ae_name according to the order we'd like in the table
  mutate(ae_name = factor(ae_name, 
                          levels = c("ae_serious_28", 
                                     "ae_ihps_28",
                                     "ae_death_28", "ae_hosp_28"), 
                          labels = c("Any serious adverse event", 
                                     "Infantile hypertrophic pyloric stenosis (IHPS)",
                                     "Mortality within 28 days of treatment",
                                     "Hospitalization within 28 days of treatment")
                          )
         )

#--------------------------
# summarize risk and risk
# difference between arms
# estimate the 95% CI using
# a linear binomial model
#--------------------------
ae_rd_ests <- foreach(aei = levels(d_ael$ae_name),.combine = rbind) %do% {
  # subset to the particular AE
  di <- d_ael %>% filter(ae_name == aei)
  # estimate Ns by arm
  dns <- di %>% 
    group_by(tx) %>%
    summarize(N = n(), n = sum(ae, na.rm = TRUE), .groups = "keep")
  
  # return results
  
  ## Bootstrap the RD for IHPS to get 95% CIs because it is so rare
  # note: this bootstrap intentionally does not exclude missing values
  # in the denominator and estimate of the risk difference
  # because IHPS was considered to be measured for all children in the trial
  if (aei == "Infantile hypertrophic pyloric stenosis (IHPS)") {
    set.seed(435352)
    aeboot <- foreach(booti = 1:1000, .combine = rbind) %do% {
         dii <- di[sample(1:nrow(di),replace=TRUE),]
         ia_N <- length(dii$ae[dii$tx=="Azithromycin"])
         ia_n <- sum(dii$ae[dii$tx=="Azithromycin"],na.rm=TRUE)
         ip_N <- length(dii$ae[dii$tx=="Placebo"])
         ip_n <- sum(dii$ae[dii$tx=="Placebo"],na.rm=TRUE)
         rd <- (ia_n/ia_N) - (ip_n/ip_N)
         return(rd)
    }

    pct_a <- dns$n[2]/dns$N[2]
    pct_p <- dns$n[1]/dns$N[1]
    rd <- pct_a - pct_p
    rd_ci <- quantile(aeboot,probs=c(0.025,0.975))
    res <- data.frame(ae_name = aei, 
                      N_a = dns$N[2], n_a = dns$n[2], pct_a = pct_a*100,
                      N_p = dns$N[1], n_p = dns$n[1], pct_p = pct_p*100,
                      rd, rd_se = NA_real_, rd_min95 = rd_ci[1], rd_max95 = rd_ci[2]) 
  } else {
    
  # fit a linear binomial model. coefficient on tx estimates the RD
  fiti <- glm(ae ~ tx, data = di, family = binomial(link = "identity"))
  sumi <- summary(fiti)
  
  # get the estimate and its SE
  esti <- fiti$coefficients[2]
  sei <- sqrt(diag(sumi$cov.unscaled)[2])
  res <- data.frame(ae_name = aei, 
                    N_a = dns$N[2], n_a = dns$n[2], pct_a = 100*dns$n[2]/dns$N[2],
                    N_p = dns$N[1], n_p = dns$n[1], pct_p = 100*dns$n[1]/dns$N[1],
                    rd = esti, rd_se = sei, rd_min95 = esti -1.96*sei, rd_max95 = esti + 1.96*sei)
  }
  
  res
}
```

```{r SAE summary table RDs}
#--------------------------
# print a summary table
#--------------------------
ae_rd_tab <- ae_rd_ests %>%
  mutate(npct_a = paste0(n_a," (",sprintf("%1.2f",pct_a),"%)"),
         npct_p = paste0(n_p," (",sprintf("%1.2f",pct_p),"%)"),
         rd95ci = paste0(sprintf("%1.2f",rd*100),"% (", sprintf("%1.2f",rd_min95*100),"% to ",sprintf("%1.2f",rd_max95*100),"%)")
         )

knitr::kable(ae_rd_tab %>% select(ae_name,npct_a,npct_p,rd95ci),
             align = "lrrr",
             col.names = c("Serious Adverse Event", 
                          paste0("Azithromycin", " \n N=", unique(ae_rd_ests$N_a), " n(%)"),
                           paste0("Placebo", " \n N=", unique(ae_rd_ests$N_p), " n(%)"),
                           "RD (95% CI) *"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE) %>%
  footnote(symbol = c("RD: risk difference (azithromycin - placebo), in percentage points. The 95% CI for IHPS was estimated with a non-parametric bootstrap (1000 iterations) due to the small number of events."))

```


## Differences between groups in AEs

```{r AE differences between arms}
#--------------------------
# pivot the AE symptoms to
# long format
#--------------------------
d_ael <- df_analysis %>% 
  filter(an_ae == 1) %>% 
  select(-c(ae_serious_28, ae_ihps_28, ae_death_28, ae_hosp_28)) %>% 
  group_by(childID,tx) %>%
  select(childID,tx,starts_with("ae")) %>%
  # reshape longer to make it easier to iterate the analysis
  pivot_longer(cols=starts_with("ae_"),names_to = "ae_name",values_to = "ae") %>%
  # ensure treatment assignment is a factor
  # order ae_name according to the order we'd like in the table
  mutate(ae_name = factor(ae_name, 
                          levels = c("ae_non_serious_21", "ae_vomiting_21",
                                     "ae_vomitAfterfeeding_21", 
                                     "ae_projVomit_21", 
                                     "ae_diarrhea_21", "ae_fever_21","ae_abpain_21", 
                                     "ae_rash_21", "ae_constip_21"),
                          labels = c("Any non-serious adverse event", "Vomiting (Any)", "Vomiting after every feed", 
                                     "Projectile vomiting", "Diarrhea", "Fever", "Abdominal Pain", "Rash", "Constipation")))

#--------------------------
# summarize risk and risk
# difference between arms
# estimate the 95% CI using
# a linear binomial model
#--------------------------
ae_rd_ests <- foreach(aei = levels(d_ael$ae_name),.combine = rbind) %do% {
  # subset to the particular AE
  di <- d_ael %>% filter(ae_name == aei)
  # estimate Ns by arm
  dns <- di %>% 
    group_by(tx) %>%
    summarize(N = n(), n = sum(ae, na.rm = TRUE), .groups = "keep")
  
  # return results
  
  ## Bootstrap the RD for projectile vomiting to get 95% CIs because it is so rare
  if (aei == "Projectile vomiting") {
    set.seed(324324)
    aeboot <- foreach(booti = 1:1000, .combine = rbind) %do% {
         dii <- di[sample(1:nrow(di),replace=TRUE),]
         ia_N <- length(dii$ae[dii$tx=="Azithromycin" & !is.na(dii$ae)])
         ia_n <- sum(dii$ae[dii$tx=="Azithromycin"],na.rm=TRUE)
         ip_N <- length(dii$ae[dii$tx=="Placebo" & !is.na(dii$ae)])
         ip_n <- sum(dii$ae[dii$tx=="Placebo"],na.rm=TRUE)
         rd <- (ia_n/ia_N) - (ip_n/ip_N)
         return(rd)
    }

    pct_a <- dns$n[2]/dns$N[2]
    pct_p <- dns$n[1]/dns$N[1]
    rd <- pct_a - pct_p
    rd_ci <- quantile(aeboot,probs=c(0.025,0.975))
    res <- data.frame(ae_name = aei, 
                      N_a = dns$N[2], n_a = dns$n[2], pct_a = pct_a*100,
                      N_p = dns$N[1], n_p = dns$n[1], pct_p = pct_p*100,
                      rd, rd_se = NA_real_, rd_min95 = rd_ci[1], rd_max95 = rd_ci[2] ) 

  } else {
    
  # fit a linear binomial model. coefficient on tx estimates the RD
  fiti <- glm(ae ~ tx, data = di, family = binomial(link = "identity"))
  sumi <- summary(fiti)
  
  # get the estimate and its SE
  esti <- fiti$coefficients[2]
  sei <- sqrt(diag(sumi$cov.unscaled)[2])
  res <- data.frame(ae_name = aei, 
                    N_a = dns$N[2], n_a = dns$n[2], pct_a = 100*dns$n[2]/dns$N[2],
                    N_p = dns$N[1], n_p = dns$n[1], pct_p = 100*dns$n[1]/dns$N[1],
                    rd = esti, rd_se = sei, rd_min95 = esti -1.96*sei, rd_max95 = esti + 1.96*sei)
  }
  
  res
}
```

```{r AE summary table RDs}
#--------------------------
# print a summary table
#--------------------------
ae_rd_tab <- ae_rd_ests %>%
  mutate(npct_a = paste0(n_a," (",sprintf("%1.2f",pct_a),"%)"),
         npct_p = paste0(n_p," (",sprintf("%1.2f",pct_p),"%)"),
         rd95ci = paste0(sprintf("%1.2f",rd*100),"% (", sprintf("%1.2f",rd_min95*100),"% to ",sprintf("%1.2f",rd_max95*100),"%)")
         )

knitr::kable(ae_rd_tab %>% select(ae_name,npct_a,npct_p,rd95ci),
             align = "lrrr",
             col.names = c("Adverse event", 
                           paste0("Azithromycin", " \n N=", unique(ae_rd_ests$N_a), " n(%)"),
                           paste0("Placebo", " \n N=", unique(ae_rd_ests$N_p), " n(%)"),
                           "RD (95% CI) *"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = "striped") %>%
  footnote(symbol = c("RD: risk difference (azithromycin - placebo), in percentage points. The 95% CI for projectile vomiting was estimated with a non-parametric bootstrap (1000 iterations) due to the small number of events."))

```



# Session Info
```{r session info}
sessionInfo()
```

