---
title: "NAITRE Trial Mortality Analysis"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Fanice Nyatigo, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---


# Summary

This report includes unmasked results from the NAITRE trial ([NCT03682653](https://clinicaltrials.gov/ct2/show/NCT03682653)). 

The report summarizes mortality outcomes in the trial.

Should you have any questions regarding this report, please write Dr. Benjamin Arnold (ben.arnold@ucsf.edu).

# Preamble
```{r preamble, message=FALSE}
#------------------------------
# load configuration file
# includes packages and masked
# assignments
#------------------------------
library(here)
source(here("NAITRE-primary/R","NAITRE-primary-Config.R"))

#--------------------------------
# set up for parallel computing
#--------------------------------
registerDoParallel(cores = detectCores() - 1)

```

# Load the analysis data

```{r read in clean data} 
#------------------------------
# read in the clean data
# rename public ID to childID
#------------------------------

# primary analysis file
df_analysis <- read_rds(here("NAITRE-primary/data/final","NAITRE_primary_analysis_public.rds")) %>% 
  rename(childID = id_public)

```


# Effect on mortality by 6 months

Cumulative incidence of mortality by 6 months is the trial's primary endpoint.


### Vital status at 6-months

```{r vital status at six months}
#-------------------------------
# use the final child status
# by 6 months dataset assembled
# above for the CONSORT flow
#
# reverse the tx ordering
# for table1
#-------------------------------
df_analysis %>% 
  mutate(tx_table1 = factor(tx, levels = c("Azithromycin","Placebo"))
         ) %>%
  set_variable_labels(vital_180 = "Vital Status at 6 months") %>%
  table1(~ vital_180 | tx_table1, data = .)
```


### Vital status at 6-months, in-window

```{r vital status at six months in window}
#-------------------------------
# use the final child status
# by 6 months dataset restricted
# to measurements in-window
# age 6 months +/- 6 weeks
#
# reverse the tx ordering
# for table1
#-------------------------------
df_analysis %>% 
  filter(death_status_180 %in% c("Measured, in-window", "Event prior to window")) %>%
  mutate(tx_table1 = factor(tx, levels = c("Azithromycin","Placebo"))) %>%
  set_variable_labels(vital_180 = "Vital Status at 6 months among measurements in-window") %>%
  table1(~ vital_180 | tx_table1, data = .)
```

<br>The primary analysis below reports a complete case analysis that excluded children lost to follow-up at 6 months (moved or had unknown status) when estimating the hazard ratio and its 95% CI. However, all children (present and missing) were included in the permutation test and resulting _P_-value.

### Hazard ratio permutation distribution 

Manually code the permutation test for the log(HR).  This code was folded into a function, `calc_ratio_pval()`, that is used for all primary inference. Retain the original here for internal comparison.

```{r glm model estimate of the HR}
#-------------------------------
# use the final child status
# by 6 months 
#
# set children with Moved
# or Unknown status to missing
#
# They will thus be excluded
# from the HR model estimates
# and permutation test
#-------------------------------
model_df <- df_analysis %>%
  filter(death_status_180 %in% c("Measured, in-window", "Event prior to window"))

#-------------------------------
# run the model
#-------------------------------
mod_empir <- glm(death_response_180 ~ tx, data = model_df, family = binomial(link = cloglog))

# save a model summary
mod_summary <- summary(mod_empir) 

# save the log HR and its SE
coef_empir <- mod_summary$coef[2,1]
se_empir <- mod_summary$coef[2,2]

# calculate the upper and lower limits on the HR ratio (coefficient from the model)
upper_empir <- exp(coef_empir+(1.96*se_empir)) %>% round(3)
lower_empir <- exp(coef_empir-(1.96*se_empir)) %>% round(3)
coef_empir_exp <- exp(coef_empir) %>% round(3)
```

```{r generate the null permutation distribution}
#-------------------------------
# permute the treatment allocation and rerun the model to 
# estimate distribution of the model coefficient under the Null
#-------------------------------
vitalstatus <- model_df$death_response_180
tx_alloc <- model_df$tx

Npermute <- 10000
null_vec <- foreach(simi = 1:Npermute, .combine = rbind) %dopar% {
  set.seed(simi)
  
  # pull random numbers the length of treatment vector from uniform distribution 0-1
  reshuffle_vec <- runif(length(tx_alloc))
  
  # reshuffle the treatment allocation based on random numbers generated
  shuffled_tx <- tx_alloc[order(reshuffle_vec)]
 
   # subset to non-missing obs
  vitalstatusi <- vitalstatus[!is.na(vitalstatus)]
  shuffled_tx <- shuffled_tx[!is.na(vitalstatus)]
  
  # run the model on shuffled allocation
  modi <- glm(vitalstatusi ~ shuffled_tx, family = binomial(link = "cloglog"))
  
  # extract and save the coefficient
  return(modi$coefficients[2])

}

#-------------------------------
# calculate the 2-sided 
# permutation test p-value
#-------------------------------
p_value <- sum(abs(null_vec) > abs(coef_empir)) / length(null_vec)
```


Permutation distribution of the log hazard ratio under the null hypothesis. The vertical black line represents the estimated log hazard ratio using the actual treatment assignments.

```{r histogram of the null distribution, echo = FALSE, warning = FALSE, message = FALSE, fig.width = 5, fig.height = 5}
#-------------------------------
# plot the null distribution 
# with the empirical value
#-------------------------------
df <- data.frame(null_vec=null_vec)
ggplot(df, aes(x=null_vec)) + 
  geom_histogram(alpha=0.5, color=NA) +
  geom_vline(xintercept = coef_empir, color="black") +
  xlab("Null distribution of log(HR)") +
  theme_bw()
```

Permutation test two-sided p-value: `r sprintf("%1.2f",p_value)`

### Mortality comparison

Repeat the analysis, using the general functions that are used throughout the primary, secondary, and subgroup analyses for internal comparison.

```{r prepare vital status table}
#-------------------------------
# summary of children and deaths
#-------------------------------
df_inwindow180 <- df_analysis %>% 
  filter(death_status_180 %in% c("Measured, in-window", "Event prior to window"))

df_mortality <- 
  calc_ratio_pval(df = df_inwindow180,
                  var = "6-month mortality", 
                  outcome = death_response_180, 
                  dist_link = "binomial_clglg",
                  N = Npermute) %>% 
  mutate(ratio = paste0(sprintf("%1.2f",ratio), " (", 
                     sprintf("%1.2f",lower), " to ", 
                     sprintf("%1.2f",upper), ")")) %>% 
  select(-c(coef, lower, upper))

# render table
knitr::kable(df_mortality, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N Deaths", "%", "N Children",
                     "N Deaths", "%", "HR (95% CI)", "P-value*"),
             caption = "Cumulative incidence of mortality by 6 months and hazard ratio (HR) for comparison between arms in children measured within the pre-specified 225-day window, estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))
```

### Sensitivity Analysis

Conduct a sensitivity analysis that includes children who were still measured at 6 months, but were measured out of window.

This includes 1,652 additional children but none of them died by the 6-month visit. Therefore, estimates are nearly identical to the primary analysis.

```{r sensitivity analysis to include out of window children}
df_all180 <- df_analysis %>% 
  filter(death_status_180 %in% c("Measured, out-of-window", "Measured, in-window", "Event prior to window"))

df_survival <- 
  calc_ratio_pval(df = df_all180,
                  var = "6-month mortality", 
                  outcome = death_response_180, 
                  dist_link = "binomial_clglg",
                  N = Npermute) %>% 
  mutate(ratio = paste0(sprintf("%1.2f",ratio), " (", 
                     sprintf("%1.2f",lower), " to ", 
                     sprintf("%1.2f",upper), ")")) %>% 
  select(-c(coef, lower, upper))
  
# render table
knitr::kable(df_survival, 
             digits = 2,
             col.names = c("Outcome", "N Children", "N Deaths", "%", "N Children",
                     "N Deaths", "%", "HR (95% CI)", "P-value*"),
             caption = "Cumulative incidence of mortality by 6 months and hazard ratio (HR) for comparison between arms including the children measured out of the 225-day window, estimated using binomial regression with a complementary log-log link."
      ) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3, "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)"))
```

# Session Info
```{r session info}
sessionInfo()
```

