---
title: "NAITRE Trial Outcomes Analysis"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Fanice Nyatigo, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---



# Summary

This report includes unmasked results from the NAITRE trial ([NCT03682653](https://clinicaltrials.gov/ct2/show/NCT03682653)). 

The report summarizes primary and secondary outcomes in the trial.

Should you have any questions regarding this report, please write Dr. Benjamin Arnold (ben.arnold@ucsf.edu).

# Preamble
```{r preamble, message=FALSE}
#------------------------------
# load configuration file
# includes packages and masked
# assignments
#------------------------------
library(here)
source(here("NAITRE-primary/R","NAITRE-primary-Config.R"))

#--------------------------------
# set up for parallel computing
#--------------------------------
registerDoParallel(cores = detectCores() - 1)

```

# Load the analysis data

```{r read in clean data} 
#---------------------------------
# load the preprocessed data
# rename public ID to childID
#---------------------------------

# primary analysis file
df_analysis <- read_rds(here("NAITRE-primary/data/final","NAITRE_primary_analysis_public.rds")) %>% 
  rename(childID = id_public)

```

```{r subset the datasets to analysis datasets}
#-------------------------------
# subset analysis dataset 
# to relevant analysis populations
# for each outcome
#-------------------------------

# death
df_death_180 <- df_analysis %>% filter(death_status_180 %in% c("Measured, in-window", "Event prior to window"))
df_death_90 <- df_analysis %>% filter(death_status_90 %in% c("Measured, in-window", "Event prior to window"))
#df_death_neo <- df_analysis %>% filter(death_status_neo %in% c("Measured, in-window", "Event prior to window"))

# death and/or hosp
df_death_hosp <- df_analysis %>% filter(death_status_180 %in% c("Measured, in-window", "Event prior to window") |
                                          hosp_status_180 %in% c("Measured, in-window", "Event prior to window"))
# hospitalization
df_hosp_180 <- df_analysis %>% filter(hosp_status_180 %in% c("Measured, in-window", "Event prior to window"))
df_hosp_90 <- df_analysis %>% filter(hosp_status_90 %in% c("Measured, in-window", "Event prior to window"))
df_hosp_21 <- df_analysis %>% filter(hosp_status_21 %in% c("Measured, in-window"))

# any clinic visits (and reason-specific clinic visits)
df_clinic_180 <- df_analysis %>% filter(clinic_status_180 %in% c("Measured, in-window", "Event prior to window"))
df_clinic_90 <- df_analysis %>% filter(clinic_status_90 %in% c("Measured, in-window", "Event prior to window"))
df_clinic_21 <- df_analysis %>% filter(clinic_status_21 %in% c("Measured, in-window"))

# total clinic visits 
df_clinic_tot_180 <- df_analysis %>% filter(clinic_status_180 %in% c("Measured, in-window"))
df_clinic_tot_90 <- df_analysis %>% filter(clinic_status_90 %in% c("Measured, in-window"))
df_clinic_tot_21 <- df_analysis %>% filter(clinic_status_21 %in% c("Measured, in-window"))


```


# Secondary outcomes by treatment arm

```{r run the model on all outcomes, warning = F, message = F}
#-------------------------------
# set the number of permutation
# test replicates for P-values
#-------------------------------
Npermute <- 10000

#-------------------------------
# calculate the hr and pvalue 
# for all relevant binary outcomes
# and combine into one data frame
#-------------------------------

all_outcomes <- 
  rbind(
    # death
    calc_ratio_pval(df = df_analysis, var = "Mortality by age 28 days‡", outcome = death_response_28, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_death_90, var = "3-month mortality", outcome = death_response_90, dist_link = "binomial_clglg", N=Npermute),
    
    # death and/or hosp    
    calc_ratio_pval(df = df_death_hosp, var = "Death and/or hospitalization, 6 months", outcome = death_hosp_180, dist_link = "binomial_clglg", N=Npermute),
    
    # hospitalization
    calc_ratio_pval(df = df_hosp_21, var = "Hospitalization- 21 days", outcome = hosp_by_21, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_hosp_90, var = "Hospitalization- 3 months", outcome = hosp_by_90, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_hosp_180, var = "Hospitalization- 6 months", outcome = hosp_by_180, dist_link = "binomial_clglg", N=Npermute),
    
    # any clinic
    calc_ratio_pval(df = df_clinic_21, var = "Clinic, Any- 21 days", outcome = any_clinic_21, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_clinic_90, var = "Clinic, Any- 3 months", outcome = any_clinic_90, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_clinic_180, var = "Clinic, Any- 6 months", outcome = any_clinic_180, dist_link = "binomial_clglg", N=Npermute),
    
    # reasons for clinic visit
    calc_ratio_pval(df = df_clinic_180, var = "Malaria- 6month clinic", outcome = malaria_by_180, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_clinic_180, var = "Pneumonia- 6month clinic", outcome = pneumonia_by_180, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_clinic_180, var = "Diarrhea- 6month clinic", outcome = diarrhea_by_180, dist_link = "binomial_clglg", N=Npermute),
    calc_ratio_pval(df = df_clinic_180, var = "Fever§- 6month clinic", outcome = fever_by_180, dist_link = "binomial_clglg", N=Npermute),
    
    # total clinic
    calc_ratio_pval(df = df_clinic_tot_21, var = "Clinic, Total- 21 days", outcome = tot_clinic_21, dist_link = "poisson_lglnk", N=Npermute),
    calc_ratio_pval(df = df_clinic_tot_90, var = "Clinic, Total- 3 months", outcome = tot_clinic_90, dist_link = "poisson_lglnk", N=Npermute),
    calc_ratio_pval(df = df_clinic_tot_180, var = "Clinic, Total- 6 months", outcome = tot_clinic_180, dist_link = "poisson_lglnk", N=Npermute)
)

rownames(all_outcomes) <- NULL
```

## Binary outcomes
```{r binary outcome table}

# format for pretty output
binary_outcomes_tab <- all_outcomes %>% 
  filter(!var %in% c("Clinic, Total- 21 days", "Clinic, Total- 3 months", "Clinic, Total- 6 months")) %>% 
  mutate(ratio = paste0(sprintf("%1.2f",ratio), " (", 
                     sprintf("%1.2f",lower), " to ", 
                     sprintf("%1.2f",upper), ")")) %>% 
  select(-c(coef, lower, upper))

# render table
knitr::kable(binary_outcomes_tab,
             digits = 2,
             col.names = c("Outcome", "N total", "n outcome", "%", "N total", "n outcome", "%", "HR (95% CI†)", "P-value*")
      ) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3,  "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)", 
                      "CI, confidence interval",
                      "Mortality before 28 days of age", 
                      "Fever without another diagnosis"))
```

## Total count outcomes
```{r count outcome table}

# format for pretty output
count_outcomes_tab <- all_outcomes %>% 
  filter(var %in% c("Clinic, Total- 21 days", "Clinic, Total- 3 months", "Clinic, Total- 6 months")) %>% 
  mutate(ratio = paste0(sprintf("%1.2f",ratio), " (", 
                     sprintf("%1.2f",lower), " to ", 
                     sprintf("%1.2f",upper), ")")) %>% 
  select(-c(coef, lower, upper))

# render table
knitr::kable(count_outcomes_tab, 
             digits = 2,
             col.names = c("Outcome", "N total", "n outcome", "mean", "N total", "n outcome", "mean", "Means Ratio (95% CI†)", "P-value*")
      ) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3,  "Placebo" = 3, " " = 2)) %>%
  footnote(symbol = c("Permutation P-value (10,000 replicates)", 
                      "CI, confidence interval"))
```

# Subgroup analyses for the primary outcome (6-month mortality)

```{r subgroup analyses, warning = F, message = F}
# re-factor tx and response 
# to be in the order required for the contingency table
d_interaction <<- df_death_180 %>% 
  filter(vital_180 %in% c("Dead", "Alive")) %>% 
  mutate(tx = factor(tx, levels = c("Azithromycin", "Placebo")),
         vital_180 = factor(vital_180, levels = c("Dead", "Alive")))

# combine output
sub_ansys <- 
  rbind(
    calc_ratio_pval(df = df_death_180, var = "6-month mortality", outcome = death_response_180, dist_link = "binomial_clglg", N=1) %>% 
      mutate(grp = "All participants", 
             p_int = NA_real_) %>% relocate(grp),
    get_subgroup_ratio(popn = df_death_180, group_var = "age_cat", N=Npermute),
    get_subgroup_ratio(popn = df_death_180, group_var = "childSex", N=Npermute),
    get_subgroup_ratio(popn = df_death_180, group_var = "enroll_season", N=Npermute),
    get_subgroup_ratio(popn = df_death_180, group_var = "region_name", N=Npermute),
    get_subgroup_ratio(popn = df_death_180, group_var = "urbanicity", N=Npermute), 
    get_subgroup_ratio(popn = df_death_180, group_var = "breastfeeding", N=Npermute),
    get_subgroup_ratio(popn = df_death_180, group_var = "weight_cat", N=Npermute)
    ) %>% 
  mutate(p_int = round(p_int, digits = 2))

```

```{r subgroup analysis table}
# format for pretty output
sub_ansys_tab <- sub_ansys %>% 
  mutate(ratio = paste0(sprintf("%1.2f",ratio), " (", 
                     sprintf("%1.2f",lower), " to ", 
                     sprintf("%1.2f",upper), ")")) %>% 
  select(-c(var, coef, lower, upper, p_value, p_int))

# render table
knitr::kable(sub_ansys_tab, 
             digits = 2,
             col.names = c("6-month mortality", "N total", "n outcome", "%", "N total", "n outcome", "%", "HR (95% CI)*"),
             caption = "Cumulative incidence of 6-month mortality by subgroup and hazard ratio (HR) for comparison between arms."
      ) %>%
  pack_rows(paste0("By age at enrollment", " \n P for interaction † = ", sub_ansys %>% filter(var == "age_cat") %>% distinct(p_int)), 2, 4) %>% 
  pack_rows(paste0("By child's sex", " \n P for interaction † = ", sub_ansys %>% filter(var == "childSex") %>% distinct(p_int)), 5, 6) %>%
  pack_rows(paste0("By season of enrollment", " \n P for interaction † = ", sub_ansys %>% filter(var == "enroll_season") %>% distinct(p_int)), 7, 8) %>%
  pack_rows(paste0("By region", " \n P for interaction † = ", sub_ansys %>% filter(var == "region_name") %>% distinct(p_int)), 9, 13) %>%
  pack_rows(paste0("By urbanicity", " \n P for interaction † = ", sub_ansys %>% filter(var == "urbanicity") %>% distinct(p_int)), 14, 16) %>%
  pack_rows(paste0("By breastfeeding", " \n P for interaction † = ", sub_ansys %>% filter(var == "breastfeeding") %>% distinct(p_int)), 17, 19) %>%
  pack_rows(paste0("By baseline weight", " \n P for interaction † = ", sub_ansys %>% filter(var == "weight_cat") %>% distinct(p_int)), 20, 21) %>%
  kable_styling(bootstrap_options = "striped", full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Azithromycin" = 3,  "Placebo" = 3, " " = 1)) %>%
  footnote(symbol = c("Hazard Ratio (HR) and 95% confidence interval (CI)",
                      "Permutation P-value for test of homogeneity (10,000 replicates)"))
```

# Session Info
```{r session info}
sessionInfo()
```

