---
title: "NAITRE Trial Post-hoc Minimum Detectable Effect (MDE) Estimates"
subtitle: "CONFIDENTIAL. DO NOT DISTRIBUTE."
author: "Prepared by: Fanice Nyatigo, and Benjamin Arnold<br><br>F.I. Proctor Foundation, University of California, San Francisco"
date: "updated: `r Sys.time()`"
output: 
  html_document:
    theme: default
    highlight: default
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, 
                          encoding   = encoding, 
                          output_dir = "../../output"
                          )})
---



# Summary

This report includes post-hoc estimates of the minimum detectable effect (MDE) for the NAITRE trial ([NCT03682653](https://clinicaltrials.gov/ct2/show/NCT03682653)) based on the primary outcome, all cause mortality by 6 months. 

This information was requested during peer review.

Should you have any questions regarding this report, please write Dr. Benjamin Arnold (ben.arnold@ucsf.edu).

# Preamble
```{r preamble, message=FALSE}
#------------------------------
# clear workspace 
#
# no need to use the overall
# project configuration file
# for this simple calculation
#------------------------------
rm(list=ls())


```

# MDE

_How large would the reduction in mortality need to be to have at least 80% power to detect a difference between groups, given the size of the trial and the empirical probability of mortality in the placebo arm?_

In the final analysis, there were 9606 children in the azythromycin group (42 deaths) and 9684 children in the placebo group (50 deaths). 

Below, the calculations assume 9606 children per group. 

```{r mde}
#-----------------------------
# empirical mortality proportions
#-----------------------------
prob_azithro <- 42/9606
prob_placebo <- 50/9684

#-----------------------------
# estimate the MDE given
# the mortality proportion in the
# placebo, the sample size and
# 80% power and type I alpha of 5%
#-----------------------------
( mde_mortality <- power.prop.test(n=9606, p2=prob_placebo, sig.level = 0.05, power=0.8, alternative = "two.sided") )

#-----------------------------
# summarize the MDE in terms
# of the RR
#-----------------------------
( mde_rr <- mde_mortality$p1/mde_mortality$p2 )
```

The estimated MDE on the relative risk scale given the design and empirical probability of mortality in the placebo arm is **RR = `r sprintf("%1.2f",mde_rr)`.**

## Sensitivity analysis

Repeat the calculation above, assuming the larger 9684 children per group. There is no difference in the MDE.

```{r mde sensitivity}
( mde_mortality2 <- power.prop.test(n=9684, p2=prob_placebo, sig.level = 0.05, power=0.8, alternative = "two.sided") )
( mde_rr2 <- mde_mortality2$p1/mde_mortality2$p2 )
```

Assuming the larger number of children per group, the MDE is still **RR = `r sprintf("%1.2f",mde_rr2)`.**

# Sample size required given the observed effect

_How large would the trial needed to have been to have 80% power to detect the observed 15% relative reduction in mortality (RR=0.85), given the observed mortality rates?_

```{r sampsi}
( sampsi_mortality <- power.prop.test(p1=prob_azithro,p2=prob_placebo, sig.level = 0.05, power=0.8, alternative = "two.sided") )

```

Given the empirical probabilities of mortality, the trial would have needed to enroll **`r sprintf("%1.0f",round(sampsi_mortality$n))` children** _per group_ (**`r sprintf("%1.0f",round(sampsi_mortality$n)*2)` total children**) to have 80% power.


# Session Info
```{r session info}
sessionInfo()
```

